Reproducible Research Peer Assesment 1
=====================================
first we will load the data, create correct date type and remove NAs

```{r}
library(ggplot2)
library(plyr)
adf <- read.csv('activity.csv')
adf$date <- as.POSIXlt(adf[, 2])
actdf <- adf[!is.na(adf[,1]),]
```
What is the mean total number of steps per day?
===
Here is a histogram of the steps taken per day
```{r}
actsummary <- ddply(actdf, .(date), summarize, Total_Steps = sum(steps))
actsummary <- actsummary[order(actsummary$date),]

p <- ggplot(actsummary, aes(date, Total_Steps)) + geom_histogram(stat = 'identity')
p 
```
The mean number of steps per day is 
```{r}
mean(actsummary$Total_Steps)
```
And the median number of steps per day is
```{r}
median(actsummary$Total_Steps)
```

What is the average daily activity pattern?
===
Here is a plot of the average steps taken during each time interval

```{r}
q2summary <- ddply(actdf, .(interval), summarize, avestep = mean(steps))
q2summary$interval <- as.integer(q2summary$interval)
q2p <- ggplot(q2summary, aes(interval, avestep)) + geom_line()
q2p
```
The interval that takes the most steps on average is
```{r}

maxfind <- order(q2summary$avestep, decreasing = TRUE)
q2summary[maxfind[1],1]
```

Inputting Missing Values
===
The total number of records with missing values is
```{R}
sum(is.na(adf$steps))
```

We will now replace NAs with average steps for that interval
```{r}
for (i in which(sapply(adf, is.numeric))) {
    adf[is.na(adf[, i]), i] <- mean(adf[, i],  na.rm = TRUE)
}

```

Here is a histogram of the steps taken per day with NAs replaced
```{r}
asummary <- ddply(adf, .(date), summarize, Total_Steps = sum(steps))
asummary <- asummary[order(asummary$date),]

p2 <- ggplot(actsummary, aes(date, Total_Steps)) + geom_histogram(stat = 'identity')
p2 
```
The mean number of steps per day is 
```{r}
mean(asummary$Total_Steps)
```
And the median number of steps per day is
```{r}
median(asummary$Total_Steps)
```
the mean number is not affected because of the method of imputing draws from
mean steps per day.
The median increases because a day with NAs went from below the median to above,
this caused the median to be the same as the mean.


Are there differences in activity patterns between weekdays and weekends?
===

```{r}
adf$date <- as.POSIXct(adf[, 2])        
adf <- ddply(adf, .(steps,date,interval), transform, weekday = weekdays(date))
adf <- arrange(adf, weekday)
weekends <-adf[10369:14976,]
weekday <- adf[1:10368,]
```

here is the weekend plot followed by the weekday plot
```{r}
lastplot <- ddply(weekends, .(interval), summarize, avestep = mean(steps))
lastplot2 <- ddply(weekday, .(interval), summarize, avestep =mean(steps))

plotty <- ggplot(lastplot, aes(interval,avestep)) + geom_line()
plotty2 <- ggplot(lastplot2, aes(interval,avestep)) + geom_line()

```
as you can see people are much more active on the weekends.